---
title: "Karol Cieply lab4"
output: github_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

## Including Code

You can include R code in the document as follows:

```{r cars}
summary(cars)
```




## Mean

$${\displaystyle A={\frac {1}{n}}\sum _{i=1}^{n}a_{i}={\frac {a_{1}+a_{2}+\cdots +a_{n}}{n}}}$$
```{r mean}
mean((c(6,9,4,3,4,9,4,2,7,3,5,4,7,1,3,5,1,7,3,4,7,9,7,2,9,1,8,2,7,4)))
```

## Variance
<br />
$$\sigma^2 = \frac{\displaystyle\sum_{i=1}^{n}(x_i - \mu)^2} {n}$$
$$\sigma^2 = \frac{\displaystyle\sum_{i=1}^{n}(x_i - \mu)^2} {n-1}$$

```{r var}
var(c(5,3,3,2,4,4,8,9,6,1,10,8,5,2,5,2,4,10,6,3,1,9,3,4,1,9,10,7,1,7))
```

## Mean price of diamonds based on color and cut
```{r csv}
diamonds <- read.csv("diamondsCSV.csv")
diamondssum <- aggregate(diamonds$price, list(diamonds$cut), sum)
diamondsavg <- aggregate(diamonds$price, list(diamonds$cut), mean)
diamondsavgcolor <- aggregate(diamonds$price, list(diamonds$cut, diamonds$color), mean)
diamondsavgcolorre <- reshape(diamondsavgcolor, idvar = "Group.1", timevar = "Group.2", direction = "wide")
head(diamondsavgcolorre)
```

## Explaining what is outlier 
In statistics, an outlier is a data point that differs significantly from other observations.
```{r outlier}
plot(iris$Sepal.Width ~ iris$Species)
```




```{r plot}
X <- seq(-3,3,by=0.2)


plot(X,dnorm(X),type="l")

plot(X,pnorm(X),type="l")
```

